---
title: "Data Science HW-2"
author: "Kasturi Bhamidipati"
date: "2022-10-02"
output: github_document
---

First we load the libraries that we need! 
```{r load pacakges}
library(tidyverse)
library(readxl)
library(skimr)
```

# Problem 2

We want to read and clean the Mr. Trash Wheel dataset:  
```{r Mr. Trash Wheel}
mr_trash_wheel = read_excel("Data/Trash Wheel Collection Data.xlsx", sheet = 1, range = "A2:N549") %>%
  janitor::clean_names()%>%
  mutate(sports_balls = as.integer(sports_balls), wheel = "mr")
```

We want to read and clean the Prof. Trash Wheel dataset: 

```{r Prof. Trash Wheel}
prof_trash_wheel = read_excel("Data/Trash Wheel Collection Data.xlsx", sheet = 2, range = "A2:M96") %>%
  janitor::clean_names()%>%
  mutate(year = as.character(year), wheel = "prof")
```

Now we want to combine the two datasets: 

```{r Combined Trash Wheel}
combined_trash_wheel = bind_rows(mr_trash_wheel, prof_trash_wheel) %>%
  janitor::clean_names()
```

Here is a description of my combined dataset: 

`r skim(combined_trash_wheel)`

- The total number of rows are `r nrow(combined_trash_wheel)` and number of variables are `r ncol(combined_trash_wheel)`

- There are three character variables, eleven numeric variables and one POSIXct variable which is for the date. 

- The total weight of trash collected by Professor Trash Wheel is `r sum(prof_trash_wheel$weight_tons, na.rm = TRUE)`tons

- The total number of sports balls collected by Mr. Trash Wheel in 2020 was `r sum(combined_trash_wheel[which(combined_trash_wheel$year =="2020" & combined_trash_wheel$wheel =="mr"), "sports_balls"])`
